ManyEcoEvo: an R package for working with the "Same Data, Different Analysts" Project in Ecology and Evolutionary Biology, and other many-analyst style data

### 2023-11-10 manuscript/manuscript/manuscript.Rmd
Complete statement of need section

@@

We acknowledge the contributions of all 246 analysts and co-authors of Gould et al. (2023).
EG is supported by an Australian Government Research Training Program Scholarship.

>>>>

The last ten years has seen the growing emergence of many-analyst projects, [e.g. @silberzahn2018; @botvinik-nezer2020; @huntington-klein2021; @schweinsberg2021; @breznau2022; @hoogeveen2022many; @coretta2023], predominantly in the disciplines of psychology and social and behavioural sciences, however we expect more of these types of studies to be published into the near future across other disciplines.
Moreover, collaborative and distributed big team open science projects are becoming increasingly common in ecology and evolutionary biology (e.g. NutNet <https://nutnet.org/index.php/>, DragNet, ManyBirds <http://themanybirds.com>) as well as other disciplines (ManyBabies **CITE**, ManyPrimates **CITE**) which pose computationally analogous data-manipulation and analysis problems to the many-analyst workflow addressed by the *ManyEcoEvo* approach.
While *ManyEcoEvo* was designed for the ManyEcoEvo project, the structured approach it employs establishes a workflow that can be generalised and applied to new many-analyst style data, facilitating replication of Gould *et al.* [-@gould2023], in addition to computational reproduction [@Stoudt2021], as well as new analyses of many-analyst type studies.

We are not aware of any R packages or software in other languages exclusively focussed on facilitating many-analyst style studies.
Existing many-analyst code if shared, is written as R scripts particular to that study's analysis (e.g. <https://github.com/manybabies/mb1-analysis-public>, <https://github.com/many-speech-analyses/many_analyses>).
However, there are a range of existing packages with select functionality that addresses various components of many-analyst style analyses, namely meta-analysis, its associated preparation and visualisation, as well as the visualisation component of [specification curve analysis] / [multiverse analysis] ([@Simonsohn2020], [@Steegen2016]).
Following the collection of results from teams conducting independent analyses, the data must be 1.
cleaned and prepared for 2.
meta-analysis and other model analysis, which is then 3.
subjected to model checking, statistic extraction and visualisation [@aczel2021].

We provide a suite of miscellaneous functions for data cleaning and processing tasks that we expect to be common to the sorts of data collected by many-analyst style studies.
This was the most time and effort-intensive aspect of the analysis in Gould *et al.* [-@gould2023], and we hope these functions will ease the burden on future researchers conducting many-analyst studies.

*ManyEcoEvo* provides custom functions to calculate standardize effect sizes in the form of Fishers' Z correlation scores in preparation for meta-analysis.
While some existing functions exist within other packages (e.g. `DescTools::FishersZ()`, `DiagTest3Grp::FisherZ.var()`), the method for calculating variance of the standardized effect size in Gould *et al.* [-@gould2023] differed slightly see `INSERT FUNCTION NAME`, and we wished to avoid reliance on too many external packages (Wickham & Bryan, 2023: **Chapter 10.1**).
Should *ManyEcoEvo* users wish to deviate from the effect-size standardisation method used in Gould *et al.* [-@gould2023], they are free to use their own functions provided by other software or custom functions (See **Harrer et al. 2022** chapter 3 for a range of R snippets that calculate various effect sizes).

`meta` (**CITE**) and `metafor` (**cite**) R packages are the two most commonly used softwares for undertaking meta-analysis in R [@lortie2020].
We provide functions for fitting the exact meta-analysis models used in Gould *et al.* [-@gould2023], which were implemented with the *metafor* package, however *ManyEcoEvo* users are free to supply alternative model-fitting functions should they wish to deviate from these specified models.
After model-fitting and analysis, model checking and extraction of model fit and heterogeneity statistics is implemented using a range of existing packages, including: `performance::`, `broom.mixed::`, `metafor::`.
*ManyEcoEvo* also draws on existing visualisation packages to implement forest plots (`ggforest::)`, funnel plots (`metafor::`), and s [specification curve analysis] plots to examine the influence of analysis specifications on variation in analysis outcomes (`specr::`).
Select functions from the above packages have been thoughtfully abstracted and coded into wrapper functions according to a principled and structured workflow (described below) for ease of application and generalisation to alternative datasets.

@@

We acknowledge the contributions of all 246 analysts and co-authors of Gould *et al. [-@gould2023]* please see the original manuscript for a full list of contributors to the ManyEcoEvo project.

EG is supported by an Australian Government Research Training Program Scholarship.
